Thermal tuning of phase-matching in a multimodal SOI waveguide with $\chi^{(3)}$ non linearity

### 2015-09-26 latex/davide_bazzanella.tex
final version?

\begin{figure}[ht]
	\centering
	\includegraphics[width=1\textwidth]{geometry.pdf}
	\label{fig_wg_Section}
	\caption{Section of the rectangular waveguide}
\end{figure}

@@

The new geometry is represented in Figure (\ref{fig_geometry_asym}).
\begin{figure}[!ht]
	\centering
	\includegraphics[width=.9\textwidth]{geometryA.pdf}
	\caption{Scheme of the side heater configuration.}
	\label{fig_geometry_asym}
\end{figure}

@@

\begin{figure}[!ht]
	\centering
	\includegraphics[width=.75\textwidth]{variation.png}
	\caption{Percentage variation of $n_{eff}$ for TE modes at \SI{1.55}{\um} from its value at ambient temperature in the symmetric configuration.}
	\label{fig_variation}
\end{figure}

@@

In Figure (\ref{fig_variation}) is represented the change of effective index, for wavelength $\lambda = \SI{1.55}{\um}$, from ambient temperature to \SI{400}{\K} over it in ratio to its value at $T_{amb}=\SI{293.15}{\K}$.
This curve suggest us that low orders are more influenced by the difference in temperature than high orders.
We can therefore hypothesize that configuration with a high difference between their orders could be controlled better.
In this view, as next step in enhancing the influence of temperature on effective index, different core geometries should be studied.
%, in order to maximise the effect of the overlap integrals between the optical mode and the thermal profile.
This could greatly condition the wave propagation, and therefore the thermal tuning, but it is also a difficult problem with many degrees of freedom.

>>>>

\begin{figure}[ht]
	\centering
	\includegraphics[width=1\textwidth]{geometryR.pdf}
	\label{fig_wg_Section}
	\caption{Section of the rectangular waveguide}
\end{figure}

@@

The new geometry is represented in Figure (\ref{fig_geometry_asym}).
\begin{figure}[!ht]
	\centering
	\includegraphics[width=.9\textwidth]{geometryAR.pdf}
	\caption{Scheme of the side heater configuration.}
	\label{fig_geometry_asym}
\end{figure}

@@

\begin{figure}[!ht]
	\centering
	\includegraphics[width=.7\textwidth]{effectiveTOC.pdf}
	\caption{Effective TOC for TE modes at \SI{1.55}{\um}.}
	\label{fig_variation}
\end{figure}

@@

In Figure (\ref{fig_variation}) is represented the effective TOC of different modes, for wavelength $\lambda = \SI{1.55}{\um}$.
This curve suggest us that high orders are more influenced by the difference in temperature than low orders.
We can therefore hypothesize that configuration with a high difference between their orders could be controlled better.
In this view, as next step in enhancing the influence of temperature on effective index, different core geometries should be studied.
%, in order to maximise the effect of the overlap integrals between the optical mode and the thermal profile.
This could greatly condition the wave propagation, and therefore the thermal tuning, but it is also a difficult problem with many degrees of freedom.

### 2015-09-26 presentazione/presentazione.tex
final version?

\section{Nonlinear Optics}
\begin{frame}{Nonlinear Optics}
	Frequency convertion employs nonlinear properties of materials
	\begin{equation}
		\textbf{P} = \varepsilon_0 \left( \chi^{(1)} \cdot \textbf{E} + \chi^{(2)} : \textbf{E}^2 + \chi^{(3)} \vdots \textbf{E}^3 + \cdots \right) = \mathrm{\textbf{P}_L} + \mathrm{\textbf{P}_{NL}}
		\label{eq_P_general}
	\end{equation}
	\begin{wrapfigure}{r}{0.33\textwidth}
		\centering
		\includegraphics[width=.3\textwidth]{upconverter_HITACHI.png}
	\end{wrapfigure}
	\begin{itemize}
	\item	$\chi^{(1)}$	:	linear part
	\item	$\chi^{(2)}$	:	2$^{nd}$ order nonlinearity\\ \hspace{25pt} (SHG, TWM)
	\item	$\chi^{(3)}$	:	3$^{rd}$ order nonlinearity\\ \hspace{25pt} (THG, FWM)
	\end{itemize}
\end{frame}
%\subsection{Four-Wave Mixing}
\begin{frame}{Four-Wave Mixing}
	%\begin{wrapfigure}{r}{0.22\textwidth}
	%    \includegraphics[width=.22\textwidth]{photons.png}
	%\end{wrapfigure}
	\begin{wrapfigure}{r}{0.35\textwidth}
		\vspace{-20pt}
	    \includegraphics[width=.33\textwidth]{W&K.png}
	\end{wrapfigure}
	FWM uses 3$^{rd}$ order nonlinearity $\chi^{(3)}$.
	%\begin{wrapfigure}{r}{0.22\textwidth}
	%    \includegraphics[width=.22\textwidth]{momentum.png}
	%\end{wrapfigure}

@@

	\vspace{5pt}
	Special case:\\ $\omega_1 = \omega_2 = \omega_P$, with stimulation.
\end{frame}
%\subsection{Conservation of energy and momentum}
\begin{frame}{Conservation of energy and momentum}
	\begin{figure}
		\centering
		\vspace{10pt}
		\includegraphics[width=.8\textwidth]{setup_b&w.png}
	\end{figure}
	Conservation of energy is naturally verified
	\begin{equation}
		\omega_I = 2\omega_P - \omega_S = \omega_P \pm \Omega
		\label{eq_energy}
	\end{equation}
	Conservation of momentum is imposed on the phase-mismatch
	\begin{equation}
		\Delta k = 2\,k_P(\omega_P) - k_S(\omega_S) -k_I(\omega_I) = 0
		\label{eq_momentum}
	\end{equation}
	%\begin{equation*}
	%\mathrm{where} \qquad k = \frac{\omega}{c} = \frac{2\pi}{\lambda}
	%\end{equation*}
\end{frame}
\begin{frame}
	Generation happens also for $\Delta k \approx 0$, but at lower efficiency.
	\begin{wrapfigure}{r}{0.5\textwidth}
		\vspace{10pt}
		\centering
	    \includegraphics[width=.5\textwidth]{sinc2.png}
	\end{wrapfigure}
	\begin{equation}
	\begin{split}
		I_I &\propto \left| \int_0^L exp(\mathrm{i}\Delta k z)dz\right| \\
			&\propto L^2 sinc(\Delta k / 2\pi)
	\end{split}
	\end{equation}
	We define the coherence length
	\begin{equation}
		L_{coh} = \frac{2\pi}{|\Delta k|}
	\label{eq_coherence_length}
	\end{equation}
\end{frame}
\begin{frame}
	Bulk silicon does not permit to verify both equations.
	\begin{equation}
	k(\omega) = k_0 n(\omega)
	\label{eq_k(w)}
	\end{equation}
	\begin{figure}
		\centering
%		\vspace{-5pt}
		\includegraphics[width=0.8\textwidth]{RefractiveIndexINFO.pdf}
		\label{fig_silicon_dispersion}
	\end{figure}
\end{frame}
%\subsubsection{Multimodal Waveguides}
\begin{frame}{Multimodal Waveguides}
	In silicon waveguides both equations can be verified.\\
	Therefore Eq. \eqref{eq_momentum} becomes:
	\begin{equation}
		\Delta k = 2\,\beta_P(\omega_P) -\beta_S(\omega_S) -\beta_I(\omega_I)
		\label{eq_beta}
	\end{equation}
	\begin{equation*}
		\mathrm{where} \qquad	\beta(\omega) = k_0 n_{eff}(\omega)
	\end{equation*}
	\vspace{-5pt}
	\begin{itemize}
	\item	If the waveguide supports only one mode, phase-matching is only near the frequency of the pump and $\Omega \ll \omega_P$
	\item	If the waveguide is multimodal, there are more matching conditions and also $\Omega \lesssim \omega_P$.
	\end{itemize}
\end{frame}

%\subsection{Phase-matching tuning}
\begin{frame}{Phase-matching tuning}
	\begin{wrapfigure}{r}{0.5\textwidth}
		\vspace{-10pt}
		\centering
	    \includegraphics[width=.55\textwidth]{narrow_bandwidth.png}
	\end{wrapfigure}
	Phase-matching conditions have narrow bandwidth.

@@

	\vspace{10pt}
	Different modes react differently at the change of temperature.
	\begin{equation}
		n_{eff} = n_{eff}(\omega,T_{amb}) + \mathrm{TOC}_{eff}\cdot(T_H-T_{amb})
		\label{eq_effective_index}
	\end{equation}
\end{frame}
\begin{frame}
	\begin{figure}
		\centering
		\vspace{-10pt}
		\includegraphics[width=.8\textwidth]{effectiveTOC.pdf}
		\label{fig_TOC_effective}
		\vspace{-10pt}
	\end{figure}
	It should be possible to control phase-matching with temperature.
\end{frame}
\section{Computational model}
\begin{frame}{Computational model}
	\begin{figure}
%		\vspace{-15pt}
		\centering
		\includegraphics[width=\textwidth]{geometry.pdf}
	\end{figure}
\end{frame}
%\subsection{Temperature simulation}
\begin{frame}{Temperature simulation}
	\begin{figure}
%		\vspace{-15pt}
		\centering
		\includegraphics[width=\textwidth]{thermal.pdf}
	\end{figure}
\end{frame}
%\subsection{Mode simulation}
\begin{frame}{Mode simulation}
	\begin{figure}
%		\vspace{-15pt}
		\centering
		\includegraphics[width=\textwidth]{output_modeTE.pdf}
	\end{figure}
\end{frame}
\begin{frame}{Mode simulation}
	\begin{figure}
%		\vspace{-15pt}
		\centering
		\includegraphics[width=\textwidth]{output_modeTM.pdf}
	\end{figure}
\end{frame}
\section{Data analysis}
\begin{frame}{Data analysis}
	\begin{itemize}
	\item	Data extraction from comsol to matlab.
	\item	Data classification (TE/TM, order).
	\item	Data processing:\\
			\begin{itemize}
			\item	polynomial fit of $\lambda$ and $T_H$.
			\item	evaluation of phase-mismatch \eqref{eq_momentum} with \eqref{eq_beta} and \eqref{eq_effective_index}
					\begin{equation}
					\Delta k = 2	 \frac{2\pi}{ \lambda_P } n_{eff}(\lambda_P,T_H)
								-\frac{2\pi}{ \lambda_S } n_{eff}(\lambda_S,T_H)
								-\frac{2\pi}{ \lambda_I } n_{eff}(\lambda_I,T_H)
					\end{equation}
					and coherence length $L_{coh} = \frac{2\pi}{|\Delta k|}$.
			\end{itemize}
	\item	Data selection: verify condition $L_{coh}/L_{sam} \geq 1$.
	\item	Data aggregation.
	\end{itemize}
\end{frame}
\section{Results}
\begin{frame}{Results: symmetric configuration}
	\begin{figure}
%		\vspace{-15pt}
		\centering
		\includegraphics[height=.8\textheight]{ppaat22.png}
	\end{figure}
\end{frame}
\begin{frame}
	\begin{figure}
%		\vspace{-15pt}
		\centering
		\includegraphics[height=.9\textheight]{baat2.png}
	\end{figure}
\end{frame}
\begin{frame}
	\begin{figure}
%		\vspace{-15pt}
		\centering
		\includegraphics[height=.9\textheight]{ppc2.png}
	\end{figure}
\end{frame}
\begin{frame}
	Combination \texttt{TE 517}, symmetric configuration.
	\begin{figure}
%		\vspace{-15pt}
		\centering
		\includegraphics[width=.8\textwidth]{bandwidth_movement_TE517_sym.png}
	\end{figure}
\end{frame}
\begin{frame}{Asymmetric configuration}
	\begin{figure}
%		\vspace{-15pt}
		\centering
		\includegraphics[width=\textwidth]{geometryA.pdf}
	\end{figure}
\end{frame}
\begin{frame}
	\begin{figure}
%		\vspace{-15pt}
		\centering
		\includegraphics[height=.85\textheight]{asym_ppc2.png}
	\end{figure}
\end{frame}
\section{Conclusion}
%\subsection{Improvements}
\begin{frame}{Improvements}
	\begin{itemize}
		\item	The heater should be simulated as a metal conductor media, instead of a simple boundary condition (TM modes).
		\item	Thermo-Optic Coefficient also depends on the wavelength of the signal.
		\item	Different core geometries should be studied.
		\item	Develop the correlation also between temperature and efficiency of energy conversion.
	\end{itemize}
\end{frame}
%\subsection{Summary}
\begin{frame}{Summary}
	\begin{itemize}
		\item Configurations with a high difference between their orders are to be preferred.
		\item Limited interaction between the temperature of material and different modes is a constraint.
		\item Thermal tuning of the FWM phase-match is, in fact, possible.
	\end{itemize}
\end{frame}
\end{document}

>>>>

\section{Nonlinear Optics}
\begin{frame}{Nonlinear Optics}
	Frequency convertion employs nonlinear properties of materials
	\begin{equation}
		\textbf{P} = \varepsilon_0 \left( \chi^{(1)} \cdot \textbf{E} + \chi^{(2)} : \textbf{E}^2 + \chi^{(3)} \vdots \textbf{E}^3 + \cdots \right) = \mathrm{\textbf{P}_L} + \mathrm{\textbf{P}_{NL}}
		\label{eq_P_general}
	\end{equation}
	\begin{wrapfigure}{r}{0.48\textwidth}
		\centering
		\includegraphics[width=.5\textwidth]{ppln-ovens2.jpg}
	\end{wrapfigure}
	\begin{itemize}
	\item	$\chi^{(1)}$	:	linear part
	\item	$\chi^{(2)}$	:	2$^{nd}$ order nonlinearity\\ \hspace{25pt} (SHG, TWM)
	\item	$\chi^{(3)}$	:	3$^{rd}$ order nonlinearity\\ \hspace{25pt} (THG, FWM)
	\end{itemize}
\end{frame}
%\subsection{Four-Wave Mixing}
\begin{frame}{Four-Wave Mixing}
	\begin{wrapfigure}{r}{0.33\textwidth}
		\vspace{-20pt}
	    \includegraphics[width=.33\textwidth]{photons.png}
	\end{wrapfigure}
%	\begin{wrapfigure}{r}{0.35\textwidth}
%		\vspace{-20pt}
%	    \includegraphics[width=.33\textwidth]{W&K.png}
%	\end{wrapfigure}
	FWM uses 3$^{rd}$ order nonlinearity $\chi^{(3)}$.
	%\begin{wrapfigure}{r}{0.22\textwidth}
	%    \includegraphics[width=.22\textwidth]{momentum.png}
	%\end{wrapfigure}
	\begin{wrapfigure}{r}{0.45\textwidth}
		\vspace{-55pt}
	    \includegraphics[width=.45\textwidth]{frequency.png}
	\end{wrapfigure}

@@

	\vspace{5pt}
	Special case:\\ $\omega_1 = \omega_2 = \omega_P$, with stimulation.
\end{frame}
%\subsection{Conservation of energy and momentum}
\begin{frame}{Conservation of energy and momentum}
	\begin{figure}
		\centering
		\vspace{10pt}
		\includegraphics[width=.8\textwidth]{setup_b&w.png}
	\end{figure}
	Conservation of energy is naturally verified
	\begin{equation}
		\omega_I = 2\omega_P - \omega_S = \omega_P \pm \Omega
		\label{eq_energy}
	\end{equation}
	Conservation of momentum is imposed in the phase-mismatch
	\begin{equation}
		\Delta k = 2\,k_P(\omega_P) - k_S(\omega_S) -k_I(\omega_I) = 0
		\label{eq_momentum}
	\end{equation}
	%\begin{equation*}
	%\mathrm{where} \qquad k = \frac{\omega}{c} = \frac{2\pi}{\lambda}
	%\end{equation*}
\end{frame}
\begin{frame}
	Generation happens also for $\Delta k \approx 0$, but at lower efficiency.
	\begin{wrapfigure}{r}{0.5\textwidth}
		\vspace{10pt}
		\centering
	    \includegraphics[width=.5\textwidth]{sinc2.png}
	\end{wrapfigure}
	\begin{equation}
	\begin{split}
		I_I &\propto \left| \int_0^L exp(\mathrm{i}\Delta k z)dz\right| \\
			&\propto L^2 sinc^2(\Delta k L/ 2\pi)
	\end{split}
	\end{equation}
	We define the coherence length
	\begin{equation}
		L_{coh} = \frac{2\pi}{|\Delta k|}
	\label{eq_coherence_length}
	\end{equation}
\end{frame}
\begin{frame}
	Bulk silicon does not permit to verify both equations.
	\begin{equation}
	k(\omega) = k_0 n(\omega)
	\label{eq_k(w)}
	\end{equation}
	\begin{figure}
		\centering
%		\vspace{-5pt}
		\includegraphics[width=0.8\textwidth]{RefractiveIndexINFO.pdf}
		\label{fig_silicon_dispersion}
	\end{figure}
\end{frame}
%\subsubsection{Multimodal Waveguides}
\begin{frame}{Multimodal Waveguides}
	In silicon waveguides both equations can be verified.\\
	Therefore Eq. \eqref{eq_momentum} becomes:
	\begin{equation}
		\Delta k = 2\,\beta_P(\omega_P) -\beta_S(\omega_S) -\beta_I(\omega_I)
		\label{eq_beta}
	\end{equation}
	\begin{equation*}
		\mathrm{where} \qquad	\beta(\omega) = k_0 n_{eff}(\omega)
	\end{equation*}
	\vspace{-5pt}
	\begin{itemize}
	\item	If the waveguide supports only one mode, phase-matching is only near the frequency of the pump and $\Omega \ll \omega_P$
	\item	If the waveguide is multimodal, there are more matching conditions and also $\Omega \lesssim \omega_P$.
	\end{itemize}
\end{frame}

%\subsection{Phase-matching tuning}
\begin{frame}{Phase-matching tuning}
	\begin{wrapfigure}{r}{0.55\textwidth}
%		\vspace{-10pt}
		\centering
	    \includegraphics[width=.6\textwidth]{narrow_bandwidth.png}
	\end{wrapfigure}
	Phase-matching conditions have narrow bandwidth.

@@

	\vspace{10pt}
	Different modes react differently at the change of temperature.
	\begin{equation}
		n_{eff} = n_{eff}(\omega,T_{amb}) + \mathrm{TOC}_{eff}\cdot(T_H-T_{amb})
		\label{eq_effective_index}
	\end{equation}
\end{frame}
\begin{frame}
	\begin{figure}
		\centering
		\vspace{-10pt}
		\includegraphics[width=.8\textwidth]{effectiveTOC.pdf}
		\label{fig_TOC_effective}
		\vspace{-10pt}
	\end{figure}
	It should be possible to control phase-matching with temperature.
\end{frame}
\section{Computational model}
\begin{frame}{Computational model}
	Definition of the problem: geometry, materials.\\
	Simulation with FEM in Comsol.
	\begin{figure}
%		\vspace{-15pt}
		\centering
		\includegraphics[width=\textwidth]{geometryR.pdf}
	\end{figure}
\end{frame}
%\subsection{Temperature simulation}
\begin{frame}{Temperature simulation}
	Stationary solution of heat transfer equation:\\
	parameter $T_H$ and boundary conditions.
	\begin{figure}
%		\vspace{-15pt}
		\centering
		\includegraphics[width=\textwidth]{thermal.pdf}
	\end{figure}
\end{frame}
%\subsection{Mode simulation}
\begin{frame}{Mode simulation}
	Modes as solutions of the Helmholtz equation:\\
	frequency as parameter and boundary conditions.
	\begin{figure}
%		\vspace{-15pt}
		\centering
		\includegraphics[width=\textwidth]{output_modeTE.pdf}
	\end{figure}
\end{frame}
\begin{frame}{Mode simulation}
	Both TE, TM, and \textit{leaky} modes were obtained.
	\begin{figure}
%		\vspace{-15pt}
		\centering
		\includegraphics[width=\textwidth]{output_modeTM.pdf}
	\end{figure}
\end{frame}
\section{Data analysis}
\begin{frame}{Data analysis}
	\begin{itemize}
	\item	Data extraction from comsol to matlab.
	\item	Data classification (TE/TM, order).
	\item	Data processing:\\
			\begin{itemize}
			\item	polynomial fit of $\lambda$ and $T_H$.
			\item	evaluation of phase-mismatch \eqref{eq_momentum} with \eqref{eq_beta} and \eqref{eq_effective_index}
					\begin{equation}
					\Delta k = 2	 \frac{2\pi}{ \lambda_P } n_{eff}(\lambda_P,T_H)
								-\frac{2\pi}{ \lambda_S } n_{eff}(\lambda_S,T_H)
								-\frac{2\pi}{ \lambda_I } n_{eff}(\lambda_I,T_H)
					\end{equation}
					and coherence length $L_{coh} = \frac{2\pi}{|\Delta k|}$.
			\end{itemize}
	\item	Data selection: verify condition $L_{coh}/L_{sam} \geq 1$.
	\item	Data aggregation.
	\end{itemize}
\end{frame}
\section{Results}
\begin{frame}{Results: symmetric configuration}
	\begin{figure}
%		\vspace{-15pt}
		\centering
		\includegraphics[height=.8\textheight]{ppaat22.png}
	\end{figure}
\end{frame}
\begin{frame}
	\begin{figure}
%		\vspace{-15pt}
		\centering
		\includegraphics[height=.9\textheight]{baat2.png}
	\end{figure}
\end{frame}
\begin{frame}
	\begin{figure}
%		\vspace{-15pt}
		\centering
		\includegraphics[height=.9\textheight]{ppc2.png}
	\end{figure}
\end{frame}
\begin{frame}
	Combination \texttt{TE 517}, symmetric configuration.
	\begin{figure}
%		\vspace{-15pt}
		\centering
		\includegraphics[width=.8\textwidth]{bandwidth_movement_TE517_sym.png}
	\end{figure}
\end{frame}
\begin{frame}{Asymmetric configuration}
	We aimed to increase the difference of behaviour between low and high orders.
	\begin{figure}
%		\vspace{-15pt}
		\centering
		\includegraphics[width=\textwidth]{geometryAR.pdf}
	\end{figure}
\end{frame}
\begin{frame}
	\begin{figure}
%		\vspace{-15pt}
		\centering
		\includegraphics[height=.85\textheight]{asym_ppc2.png}
	\end{figure}
\end{frame}
\section{Conclusion}
%\subsection{Improvements}
\begin{frame}{Improvements}
	\begin{itemize}
		\item	The heater should be simulated as a metal conductor media, instead of a simple boundary condition (TM modes).
		\item	Thermo-Optic Coefficient also depends on the wavelength of the signal. It is important for high $\Omega$ conversions.
		\item	Different core geometries should be studied.
		\item	Develop the correlation also between temperature and efficiency of energy conversion.
	\end{itemize}
\end{frame}
%\subsection{Summary}
\begin{frame}{Summary}
	\begin{itemize}
		\item Configurations with a high difference in TOC between their orders are to be preferred.
		\item %Limited interaction between the temperature of material and different modes is a constraint.
		Optimization of the overlap between the thermal and optical spatial profiles in order to maximize the difference of effective TOC in the selected modes.
		\item Thermal tuning of the FWM phase-match is, in fact, possible.
	\end{itemize}
\end{frame}
\end{document}
